---
title: "A Tutorial on fitting Generalized Linear Models with the GLMcat Package"

author:
  - Lorena León^[Université de Montpellier, ylorenaleonv@gmail.com]
  - Jean Peyhardi^[Université de Montpellier, jean.peyhardi@umontpellier.fr]
  - Catherine Trottier^[Université de Montpellier, catherine.trottier@umontpellier.fr]

keywords:
- key
- dictionary
- word

abstract: |
  In statistical modeling, there is a wide variety of regression models for categorical responses. Yet, no software encapsulates all of these models in a standardized format. We introduce and illustrate the utility of GLMcat, the R package we developed to estimate generalized linear models implemented under the unified specification $(r, F, Z)$, where $r$ represents the ratio of probabilities (reference, cumulative, adjacent, or sequential), $F$ the cumulative distribution function for the linkage, and $Z$ the design matrix. We present the properties of the four families of models, which must be investigated when selecting the components $r$, $F$, and $Z$. The functions are user-friendly and fairly intuitive; offering the possibility to choose from a large range of models through a combination $(r, F, Z)$. Through different examples, we compare our package with VGAM and ordinal, two popular packages for implementing GLMs for categorical data.
  

output: rmarkdown::pdf_document
bibliography: bibliography_vignette.bib
vignette: >
  %\VignetteIndexEntry{A Tutorial on fitting Generalized Linear Models with the GLMcat Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Introduction to the *(r,F,Z)* methodology:

A generalized linear model is characterized by two components: 1) the\textit{ random component} that defines the conditional distribution of the response variable $y_i$ given the realization of the explanatory variables $\boldsymbol{x_i}$; 2) the \textit{systematic component} which is determined by the \textit{linear predictor} $\eta$  (that specifies the linear entry of the independent variables), and, the link function $g$  that relates the expected response and the linear predictor. 

The random component of a GLM for a categorical response with $J$ categories is the vector $(\pi_1,...,\pi_{J})$ where $\sum \pi_r = 1$ . The linear predictor  $(\eta_1,...,\eta_{J-1})$ can be written as the product of the design matrix $Z$ and the unknown parameter vector $\beta$. The link function which characterizes this model is given by the equation  $g(\pi) = Z\beta$, with $J-1$ equations $g_j = \eta_j$. @peyhardi_new proposed to write the link function as\begin{equation}
g_j = F ^ {-1} \circ r_j \Leftrightarrow r_j = F(\eta_j) \quad \quad {j = 1,2,...,J-1}
\end{equation}where $F$ is a cumulative distribution function and $r=(r_1,...,r_{J-1})$ is a transformation of the expected value vector.

In the following, we will describe in more detail the components *Z*, *F*, *r* and their modalities.

#### Ratio of probabilities *r*:

 \begin{table}[h]
        \centering
        \begin{tabular}{{lllll}}
        \hline
        \multicolumn{1}{|l|}{}           & \multicolumn{1}{l}{Cumulative} & \multicolumn{1}{l} {{Sequential}} & \multicolumn{1}{l|}{{Adjacent}} & \multicolumn{1}{l|}{{Reference}} \\
        \multicolumn{1}{|l|}{$r_j(\pi)$    \quad   \quad  } &  \multicolumn{1}{l}{$\pi_1+...+\pi_j$}       & \multicolumn{1}{l}{$\dfrac{\pi_j}{\pi_j +...+ \pi_J}$}           & \multicolumn{1}{l|}{$\dfrac{\pi_j}{\pi_j + \pi_{j+1}}$} & \multicolumn{1}{l|}{$\dfrac{\pi_j}{\pi_j + \pi_J}$}              \\ \hline
        \multicolumn{1}{|l|}{$Y$}           & 
         \multicolumn{1}{l}{} & \multicolumn{1}{l} {\centering{{ordinal}}} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{\centering {nominal}} \\ \cline{1-5} 
         
        \end{tabular}
        \end{table}

Summary of each ratio.

####  Cumulative distribution function *F*: 
The distribution functions available in `GLMcat` to fit the models are: *Logistic*, *Normal*, *Cauchit*, *Student (with any df)*, *Gompertz* and *Gumbel*.

#### Design Matriz $Z$:
It is possible to impose restrictions on the thresholds, or on the effects of the covariates, for example, for them to vary or not according to the response categories.

- Constraints on the effects: \
It is plausible that a predictor have an specific level of impact on the different categories of the response. Thus, the $J-1$ linear predictors are of the form: $\eta_j =\alpha_j + x' \delta_j$  with $\beta = (\alpha_1,..., \alpha_{J-1}, \delta_1', ..., \delta_{J-1}')$. And, its associated design matrix is:
\begin{equation}
 Z_c=\left( \begin{array}{cccccc}
             1 &  &  & x^t &  & \\ 
              & \ddots  & & & \ddots  & \\
              &  & 1 &  &  &  x^t
            \end{array} \right)_{(J- 1) \times (J -1)(1 + p)} \ 
\end{equation}
Another case, is to constrain the effects of the covariates to be the constant across the response variable categories, therefore, there is only a global effect that is not specific to the response categories, this is known as the parallelism assumption, for which the constrained space is represented by:
\begin{equation}
Z_p=
            \left( \begin{array}{cccc}
             1 &  &  & x^t  \\ 
              & \ddots  & & \vdots  \\
              &  & 1 &  x^t
            \end{array} \right)_{(J- 1) \times (J -1 + p)}
\end{equation} The first case $(Z_c)$ is named by @peyhardi_new as the \textit{complete} design, whereas the second $(Z_p)$ as the \textit{proportional} design. These two matrices are sufficient to define all the classical models. A third option is to consider both kind of effects, complete and proportional, this in known as \textit{partial proportional design}
\begin{equation}
 Z_c=\left( \begin{array}{ccccccc}
             1 &  &  & x_k^t &  & & x_l^t  \\ 
              & \ddots  & & & \ddots &  & \vdots  \\
              &  & 1 &  &  &  x_k^t &  x_l^t
            \end{array} \right)_{(J- 1) \times ((J -1)(1 + K) +  L)} \ 
\end{equation}\

- Constraints on the intercepts:\
For the particular case of the *cumulative* ratio there are two cases: *equidistant* constraint considers that the distances between adjacent intercepts are the same for all the pairs ${(j, j+1)}$, therefore we can write the intercepts as 
\begin{equation}
\alpha_j=\alpha_1+(j-1)\theta
\end{equation}
this restriction implies that only two parameters ($\alpha_1$, the first threshold, and, $\theta$ the spacing) have to be estimated regardless the number of categories, and the *symmetric* constraint, where the categories are supposed to be equally far from the central category. For an even number of response categories, there are $J/2$ parameters to estimate, and for an uneven number there are $J/2 + 1$.

All the classical models for categorical response data, can be written as an *(r,F,Z)* triplet, as examples:\

* The multinomimal model $\equiv$  $(Reference, Logistic, Complete)$\
* The odds proportional logit model $\equiv$  $(Cumulative, Logistic, Proportional)$\
* The proportional hazard model $\equiv$  $(Sequential, Gompertz, Proportional)$\
* The continuation ratio logit model $\equiv$  $(Sequential, Logistic, Complete)$\
* The adjacent logit model $\equiv$  $(Adjacent, Logistic, Complete)$\


### Fitting *(r,F,Z)* with `GLMcat` package

We used the 223 observations of the *boy's disturbed dreams* benchmark dataset drawn from a study that cross-classified boys by their age *x* and the severity of their disturbed dreams *y* [@maxwell1961analysing]. The data is available as the object `DisturbedDreams` in package `GLMcat`. 

For more information see the manual entry for the `DisturbedDreams` data: `help(DisturbedDreams)`.

```{r , include = FALSE}
library(GLMcat)
```

```{r}
data("DisturbedDreams")
summary(DisturbedDreams)
```

We will fit the model $(Reference, Logistic, Complete)$ to the `DisturbedDreams` data using the funcion `GLMcat`. We save the fitted `GLMcat` model in the object `mod1` and we print it by simply typing its name:

```{r }
mod1 <- GLMcat(formula = Level ~ Age, ratio = "reference",
  categories_order = c("Not.severe", "Severe.1", "Severe.2", "Very.severe"),
  data = DisturbedDreams, distribution = "logistic")
# mod1
```

The most common **R** functions which describe different model features are available for the objects in `GLMcat`

* The summary of the object:

```{r }
summary(mod1)
```

* The number or observations:

```{r }
nobs_glmcat(mod1)
```

* The coefficients of the model

```{r }
coef(mod1)
```

* The log-likelihood

```{r }
logLik(mod1)
```

* Information criteria

```{r }
AIC(mod1)
BIC(mod1)
```

The `predict` function is available for predictions from the results of the `GLMcat` model. We will predict the response for 3 random observations:
```{r }
# Random observations
set.seed(13)
ind <- sample(x = 1:nrow(DisturbedDreams), size  = 3)
```

* Probabilities

```{r }
predict_glmcat(mod1, data = DisturbedDreams[ind,], type = "prob")
```

* Linear predictor

```{r }
predict_glmcat(mod1, data = DisturbedDreams[ind,], type = "linear.predict_glmcator")
```

* Cumulative Probabilities

```{r }
predict_glmcat(mod1, data = DisturbedDreams[ind,], type = "cum.prob")
```

We will assume now that we are interested in making the effect of the variable predictor proportional, to that end, we type the name of the predictor variable as the input for the parameter `proportional`. The model to fit corresponds to the triplet *(Reference, Logistic, Proportional)*:

```{r }
mod2 <- GLMcat(formula = Level ~ Age, proportional = "Age",
  categories_order = c("Not.severe", "Severe.1", "Severe.2", "Very.severe"),
  data = DisturbedDreams, distribution = "logistic")
summary(mod2)
logLik(mod2)
predict_glmcat(mod2, data = DisturbedDreams[ind,], type = "prob")
```

Another variation of the reference model is obtained by changing the distribution function. Let's fit now the model *(Reference, Student (0.5), Complete)*:
```{r }
mod3 <- GLMcat(formula = Level ~ Age,
  categories_order = c("Not.severe", "Severe.1", "Severe.2", "Very.severe"),
  data = DisturbedDreams, distribution = "student", freedom_degrees = 0.5)
summary(mod3)
logLik(mod3)
predict_glmcat(mod3, data = DisturbedDreams[ind,], type = "prob")
```
